---
title: "String and Factors"
author: "Kimberly Lopez"
date: "2024-10-15"
output: github_document
---
Load libraries 
```{r}
library(rvest)
library(p8105.datasets)
library(stringr)
library(forcats)
library(rvest)
library(dplyr)
library(tidyverse)

```

# Strings and Regex

The most frequent operation involving strings is to search for an exact match. 

- use str_detect to find cases where the match exists (often useful in conjunction with filter)
use str_replace to replace an instance of a match with something else (often useful in conjunction with mutate). 

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
str_replace(string_vec, "jeff","Jeff")

```

For exact matches, you can designate matches at the beginning or end of the line. 
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
str_detect(string_vec, "i think^")
```

They all have the same i think" at the beginning of the line. To make it either start and end with "i think" use the ^ at the beginning or end of the string.

You can designate a list of characters that will count as a match.
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Pp]umpkin")
```

To get a range of letters or numbers that count as a match: use - 

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
```

The character . matches anything.
Suppose we want to detect any character with 7 and 11 with "." 

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

If your looking for a "special" charcater like: [ ], ( ), and . You have to indicate they are special using "\". However, \ is also a special character so you have to put two \\. 

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")

```

# Thoughts on Factors 


```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex
as.numeric(vec_sex)
```

To relevel based on "reference" category in an analysis, you can use the fct_relevl() function to change the underlying level structure 

```{r}
vec_sex = fct_relevel(vec_sex, "male")
vec_sex
as.numeric(vec_sex)
```

 IN GENERAL: always recode the factor variables to the levels with the order of the data. Anytime you are dealing with a factor use the package forcats 

# National Survey on Drug Use and Health data 

revisiting example from before: 
```{r echo=TRUE}
library(rvest)
library(dplyr)
library(tidyr)
```

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) |> 
  html_table() |> 
  first() |>
  slice(-1)

head(table_marj)
```

Tidy the dataset: 

- Need to split age variable by using seperate() splitting on the charcater of an open paranthesis ( ) **make sure to use \\ since ( ) is a special character**
- and some of the values in percent need to convert to numeric 
  - mutate() and str_replace() to get rid of the parenthesis 
- Go from wide format to long format using pivot_longer()
- get rid of (P Value) from column names  using select() 

```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

head(data_marj)
```
We used stringr and regular expressions a couple of times above:

- in separate, we split age and year at the open parentheses using "\\("
- we stripped out the close parenthesis in mutate
- to remove character superscripts, we replaced any character using "[a-c]$" then use "" to see if there is anything else to check if there is anything else after a-c that need to be cleaned -> R would give a warning message if there was

Now if we want to look at only data for 12-17 age group; readable plot; treating state as a factor to reorder according to the median percent value 

- use the detail axis.test.x = element_text(angle = ##, hjust = 1) to rotate the anxis labels

```{r}
data_marj |>
  filter(age == "12-17") |> 
  mutate(State = fct_reorder(State, percent)) |> 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```




